\documentclass[12pt]{article}

% House style preamble
\input{.house-style/preamble.tex}
\usepackage{float}

% Update PDF metadata
\hypersetup{
  pdftitle={Not every stable cluster is homeostatic: Stability, network order, maintenance, and control in projectible kinds}
}

\title{Not every stable cluster is homeostatic:\\ Stability, network order, maintenance, and control in projectible kinds}
\author{Brett Reynolds \orcidlink{0000-0003-0073-7195}%
\thanks{Contact: \href{mailto:brett.reynolds@humber.ca}{brett.reynolds@humber.ca}}\\
Humber Polytechnic \& University of Toronto}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Homeostatic property cluster theory gave natural-kinds realism a powerful
anti-essentialist form: a kind can be real when a clustered profile supports
induction and explanation by accommodating our classifications to causal
structure rather than by sharing an essence. In the HPC-descended debate, the
homeostatic component can be made to cover several achievements that aren't the
same:
a stable projectible profile, network order, maintenance, and corrective control.

This paper treats projectibility as the target of kind claims and separates two
counterfactuals that HPC talk can run together: what would happen if a
\term{stabilizer} were removed, and whether some process registers and answers
perturbations to the profile. The central result is separability: a profile may
be stable, network ordered, or maintained without yet being control-secured. That
result holds whatever we call the top tier. If
\term{homeostatic} is to mark a distinctive achievement, it should name only the
case where perturbation-sensitive organization preserves a specified higher-scale
relation through lower-scale variation.

The Watt governor shows how lower-scale feedback can produce higher-scale
matching; glycemic regulation shows how a regulated range can be constituted by
self-maintenance; interjection tokens show projectible stability that no
specified \term{controller} secures; and monitored agreement production shows
what a specified linguistic loop would have to provide.

The payoff is a diagnostic that keeps those achievements apart without turning
\term{homeostatic} into a loose synonym for stable or projectible.
\end{abstract}

\noindent Keywords: natural kinds; homeostatic property clusters;
projectibility; homeostasis; causal control; linguistic categories

\section{Introduction}

A quiet ambiguity runs through recent anti-essentialist kind theory. Boyd's
homeostatic property cluster framework, Slater's stable property cluster proposal,
and Khalidi's causal-network account all reject essences, and they all aim to
explain why some categories support projection while others do not. Each, though,
locates the basis for projection somewhere different.

Boyd ties projectibility to homeostatic mechanisms that maintain the cluster
\citep{boyd1991}; Slater treats natural kindness as a matter of cliquish
stability rather than of any particular maintaining mechanism \citep{slater2015};
and Khalidi locates projectibility in the
causal-network relations among core and derivative properties
\citep{khalidi2013,khalidi2018nodes}. In this journal, \textcite{onishi2022}
argue that the notion of a homeostatic mechanism in HPC theory remains
underexplicated. This paper takes that opening seriously. The title names
a pressure inside that vocabulary: even when no one explicitly identifies every
stable cluster with a homeostatic one, HPC-descended vocabulary can leave the
securing structure for a given projection underdetermined.

The looseness is clearest in the homeostatic-mechanism component of the framework.
Boyd gives that component considerable flexibility, treating the clustering as
\enquote{metaphorically (sometimes literally)} homeostatic and allowing it to be
sustained by mutual property support, underlying mechanisms, or both
\citep{boyd1999}. That flexibility is a virtue for anti-essentialist realism, but
it also makes it easy to slide from stable profile to stabilizer, from stabilizer
to homeostasis, and from homeostasis to projectible kind, as though these were one
achievement.

The order should run the other way. Start from a projection target, ask what
empirical profile supports it, ask what ordering relations make the profile
non-accidental, ask what maintains it, and only then ask whether the maintenance
is corrective control. This order keeps Boyd's projectibility frame in place
while asking that \term{homeostasis} be used in a more demanding,
control-theoretic sense. Weinberger's recent account of homeostasis as causal
control supplies the strictest standard
\citep{weinberger2026HomeostasisCausalControl}; the contribution here is to
place that standard inside an HPC-facing taxonomy of weaker securing structures.

The answer to \textcite{onishi2022} is therefore a division of evidential labour.
The original work of the paper is the \term{stabilizer}/\term{controller} split
and the intermediate taxonomy around it: a maintainer can stabilize a projectible
profile without registering and correcting perturbations to the relation that made
it projectible. Network order and maintenance are likewise distinct evidential
demands, and many real kinds may satisfy some of these demands without satisfying
others.

The cases have to show that separability. The scope is Boyd-style projectible
realist kind claims. Many classifications earn their keep by mapping variation,
stabilizing measurement, organizing historical comparison, or coordinating inquiry
before they support strong projection. Those uses are legitimate, but the
diagnostic here concerns categories offered as real kinds because they support
prediction, explanation, induction, or controlled intervention.

The paper tests the diagnostic across a machine, an organism, and two linguistic
phenomena. The linguistic payoff is sober: the diagnostic mostly tells us when not
to call a category \term{homeostatic}, while preserving a standard for future
positive cases.

\section{Projectibility first}

Boyd's framework works because projectibility doesn't require necessary and
sufficient conditions. A category can support induction when its associated
properties are connected non-accidentally, even where no single property is shared
by every member. On Boyd's accommodation thesis, successful classification is a
matter of fitting our inferential practices to causal structure, so that induction
and explanation \enquote{always require that we accommodate our categories to the
causal structure of the world} \citep[224]{boyd1991}. We construct the categories,
but whether they succeed depends on a structure we don't control. That structure
carries the explanation.

The problem of \term{projectibility} comes from Goodman's new riddle of
induction, where a predicate can fit all observed cases and still fail to license
projection to unobserved ones \citep{goodman1983}.

Goodman sets up the problem here; the diagnostic departs from his entrenchment
response. That response treats projectibility as inherited from a predicate's
history of successful projection. This diagnostic follows Boyd in asking a
different question: once a projection target is declared, what structure makes
the projection reliable, and what empirical tests constrain the declaration
against gerrymandering?

\textcite{jantzen2015} makes the same problem explicit for scientific practice by
distinguishing system induction from kind induction; this argument uses
\term{projection target} broadly enough to cover both.

Boyd's contribution was to ground projectibility in causal structure without
grounding it in essences, which is also what gave HPC theory its purchase against
the worry, which Boyd associated with Hacking, that where a kind's boundaries are
fuzzy its membership is fixed by \enquote{people in concert} rather than by nature
\citep{boyd1991}.

The same point can be put in Boyd's later terms. A programmatic definition states
the inductive or explanatory role a kind term is supposed to play; an explanatory
definition states the properties and causal structures in virtue of which it plays
that role \citep{boyd1999}. The projection target in this diagnostic is
programmatic in that sense. Stable profile, network order, maintenance, and
control are candidate explanatory supports.

One reason, then, is that it helps to begin from the projection target before
mechanism. Mechanisms matter to the theory because they explain why projection is
reliable, so a category that supports no projection isn't redeemed by having a
tidy causal story behind it, and a category that does support projection is
already doing much of the work a realist about kinds cares about. The task of the
rest of the paper is to ask, for categories that support projection, which
further evidential demands they meet.

The risk is that this question can be made too easy. If the only thing asked is
whether a cluster recurs, a great many categories will qualify, because recurrence
on its own is cheap. Stable co-occurrence may license some inference, but an
analysis still has to specify which projection is being made before the
category becomes theoretically interesting, and that happens only when its profile
supports prediction, explanation, or controlled intervention.

Projection is therefore field-relative without being arbitrary. A category is
projectible for some family of questions, contrasts, and evidential practices:
metabolic physiology, interactional linguistics, and grammar license different
expectations from different profiles. But the field doesn't make the projection
successful. Once the target is declared, the profile still has to support reliable
expectations under the relevant contexts and interventions.

Nor is the declaration self-certifying. It is a proposal about what the profile
lets us infer, and it earns its role only when the declared projection survives
empirical tests under the relevant contrasts. An analyst can redescribe a profile
in many ways, but the redescription won't secure a kind claim unless it improves
projection.

\section{Observed stability and invariance}

Slater's stable property cluster account presses the useful point that natural
kindness needn't be assessed by first identifying a homeostatic mechanism. What
matters on this view is \term{cliquish stability}: observing enough of a cluster
licenses expectations about the larger cluster across the counterfactual range
relevant to a domain of inquiry.

Slater's point is that natural kindness can be assessed through this
projectibility-shaped stability without first identifying the mechanism, if any,
by which the stability is maintained \citep{slater2015}. This yields the first
distinction the paper needs, because a stable and projectible cluster may be
perfectly real without being homeostatic in any stronger sense.

Once the target is fixed, stable-profile evidence is the first demand. It asks
whether the profile supports the declared projection: whether observing part of
the cluster licenses expectations about unobserved features or behavior across
relevant contexts. That evidence can make the category real and projectible. It
still leaves open which stronger securing claim has been earned. Stable-profile
evidence alone doesn't show network order, identify a stabilizer, or exhibit
corrective control.

The dynamical-kinds account in \textcite{jantzen2015} is a discovery-oriented
complement: it asks how projectible systems or kinds can be found by looking for
dynamical symmetries, transformations to which a causal structure is indifferent.
This diagnostic asks what has been secured once a profile is proposed.
Symmetry or invariance evidence may show stable profile or network order, but not
maintenance or perturbation-sensitive control.

This is where Onishi and Serpico's dilemma enters. If an account stops at
stable-profile evidence, it may be too slender for some explanatory projects. It
can say that the projection is supported across the relevant range, but it may
not explain why newly discovered properties should be expected in other members,
or how the same kind supports projection across context-dependent realizations.

Craver's strict-mechanism reading shows why the problem isn't solved by demanding
mechanisms: deciding which factors belong to a mechanism, when mechanisms share a
type, and where one ends introduces conventional perspective into mechanism-based
kind claims \citep{craver2009}.

Onishi and Serpico generalize the same pressure to causal-network order. If
causal structure is retained through causal-network order, as Khalidi proposes,
causal graphs still require choices about boundaries, abstraction level, and where
a network should be carved when processes and clusters fail to map one-to-one
\citep{onishi2022}. Their broader observation is that the notion of a homeostatic
mechanism has never been fully explicated, and that interest-relativity is
unlikely to be eliminable.

The diagnostic offered here doesn't try to remove that interest-relativity so
much as to make it explicit. Asking first what a category lets us project, and
against which declared expectation, fixes the explanatory target, and the question
of securing structure is then asked relative to that target rather than in the
abstract. Interest-relativity becomes a parameter the analysis declares rather than
a cost it pays without noticing. The purpose is human; whether the profile
supports projection for that purpose is an empirical constraint.

Woodward's account of invariance adds a stronger demand for explanatory
projections or projections that support intervention. Invariance is not de facto
persistence across space and time. It is a modal relation: a generalization is explanatory
when the relation it describes would continue to hold under an appropriate range
of interventions on variables in that relation \citep{woodward2001}.

The common-cause lesson matters here. A correlation may be stable across many
ordinary contexts and still fail to explain if an intervention on one effect
doesn't manipulate the other. The right intervention reveals counterfactual
dependence; the wrong one may break, bypass, or mislocate the relation being
tested.

The first diagnostic step after the projection target is correspondingly modest.
A stable property cluster should support a Slater-style projection across a
specified range of contexts, with evidence that goes beyond recurrence and shows
that observing part of the profile licenses a useful expectation about another
part. When the claimed payoff is explanation or controlled intervention, the
stronger Woodwardian test asks whether the relation that licenses the projection
is invariant under relevant what-if changes. That is not yet a control claim:
invariance is the test; corrective control is a later claim about why the
invariance holds.

\section{Network order without control}

Khalidi's network account addresses a different shortcoming. It also starts from
projectibility: natural kind predicates support reliable inferences to other
predicates, and causal-network order explains why those inferences are not
accidental. A cluster is more than a list of associated properties because its
properties can be ordered, with some core causal properties giving rise to
derivative properties in recurrent causal processes
\citep{khalidi2013,khalidi2018nodes}.

That matters because projectibility has two dimensions in Khalidi's discussion:
the variety of projections a kind supports and the generality or strictness of
those projections. Network order supplies the ontological correlate. Richer and
tighter causal relations explain richer and tighter projections, while a merely
associated bundle of features has no comparable structure.

That order is the strength of the account. The point here is narrower:
ordered dependence, even when objective and explanatory, doesn't by itself amount
to feedback-sensitive correction.

Ordered dependence is still some distance from corrective control. A node in a
causal network can support projection because the properties around it stand in
systematic causal relations. This remains compatible with the system never
tracking a departure, never allowing lower-scale variables to adjust, and never
preserving a higher-scale relation through any such adjustment. Network order can
explain projectibility without amounting to homeostasis.

The network-order step therefore asks whether the profile has a structure that
explains which projections are licensed. The relevant evidence consists of
asymmetries among the properties: which features ground or condition others,
which are derivative, which are central because many projections flow through
them, and which are only peripheral cues. The characteristic failure is the use
of causal vocabulary for a merely unordered list, since redescribing
co-occurrence in causal terms doesn't by itself establish that any ordering
obtains.

I use \term{network order} broadly across the diagnostic. This is an extension of
Khalidi's evidential role, not a claim that every ordering relation is
Khalidi-style causal naturalness. In strictly biological or mechanical cases the
ordering will usually be causal, whereas in linguistic and social cases the
relevant dependencies may instead be functional, conventional, sequential,
normative, or evidential. What matters at this diagnostic step is that the
profile has a structure that explains the declared projection.

The broad use still needs a constraint. A non-causal ordering relation should be
asymmetric: one feature conditions, licenses, constrains, or structures another.
Network order requires more than evidential usefulness. A cue is evidentially
privileged only when it tracks an ordering relation in the profile rather than
merely co-occurring with the rest of the bundle. If the analyst wants the
stronger Khalidi-style natural-kind claim, convention, evidential usefulness, or
blueprint-like co-instantiation is insufficient on its own; causal,
causal-historical, or causal-normative support has to be shown.

\section{From stabilizers to controllers}

Maintenance introduces two questions that are easy to run together. Boyd's
\term{homeostatic} label was deliberately broad: a cluster may be sustained
because some properties favor others, because underlying mechanisms maintain
their co-occurrence, or both \citep{boyd1999}. That breadth is part of what made
the framework useful.

For the diagnostic here, whatever keeps a cluster clustered can be called its
\term{stabilizer}, in the sense that removal would let the category fragment,
drift, or dissolve. The stabilizer is a functional role, and the causal structure
that fills the role is the mechanism. Maintenance, in this sense, is the umbrella
term.

The question the paper presses is which stabilizers are also controllers. A
\term{controller} is the narrower case: a stabilizer with feedback-sensitive
corrective organization that preserves a higher-scale relation through lower-scale
variation. Not every stabilizer meets this condition. Reliable co-occurrence can be
maintained by entrenchment, by transmission, or by a frozen accident that nothing
actively corrects, and ordered dependence can be maintained by causal wiring that
carries no error signal and returns the system to no target.

The two ideas answer different counterfactuals, which is why they come apart. The
stabilizer question is about the maintainer, and it asks what would happen to the
cluster if the maintaining process were removed. The controller question is about
the profile under threat, and it asks whether some maintaining process registers
and answers a perturbation that threatens the profile in a way that brings the
profile back toward a relation worth projecting from.

A case can satisfy the first condition and fail the second, since entrenchment and
transmission may hold a profile together without detecting or correcting a
departure. The space between a merely held profile and an actively corrected one is
where \term{homeostatic} tends to be overclaimed.

Read this way, maintenance is independent of network order in both directions. A
stable profile can support projection before its maintainer is known, and a
network-ordered profile can explain why projection is reliable without yet showing
what keeps the network in place. Conversely, a maintainer may keep a profile in
place without giving the profile much internal order. Once a maintainer is
identified, the further question is whether that maintainer is a controller. Only
the controller variety earns \term{homeostatic} in the strict control-theoretic
sense, and the next section says what control adds.

\section{Corrective control}

Control-secured kind claims are more demanding, and the first thing to get right
is the level at which the control is located. A controller can of course be a kind
of system, but the kind label leaves open what senses, compares, or corrects. A
candidate kind earns the \term{homeostatic} label only when its projectible
profile is maintained by a specified system or practice with feedback-sensitive
corrective organization.

That specification is itself relative to the projection target. The analyst has
to say which higher-scale relation matters, which lower-scale variables may vary,
which perturbations threaten the relation, and which response pathway preserves
it. The scale split therefore remains interest-relative, but it has to be declared
before the case is classified.

The proposed target has to appear in the corrective dynamics: the relevant
perturbation should disturb the candidate relation, the response pathway should
answer that perturbation, and the preserved relation should be the one that
supports the projection. Relabelling a maintained profile as control doesn't
supply that pattern.

Weinberger's central lesson is that homeostasis requires more than constancy,
because a controlled system can preserve a higher-scale relation precisely by
letting its lower-scale variables move. His dynamic causal models show that
feedback systems, and the Watt governor in particular, produce control relations
that static causal models and a simple vocabulary of intervention tend to obscure
\citep{weinberger2026HomeostasisCausalControl}.

In the governor case, the neglected relation is matching steam admission to
workload, not holding shaft speed fixed. That point fits a broader control
literature on organized systems. In Bich and Bechtel's terms, control mechanisms
measure variables and use those measurements to set constraints in controllers,
which then act on flexible constraints in the mechanisms they control
\citep{bichBechtel2022control}. In living systems, control is often organized
across networks of mutually dependent processes rather than issued from a single
command point \citep{bichMossioEtAl2016,bichBechtel2022organization}.

Two consequences are worth drawing out. The first is that holding a variable
rigidly fixed can be a sign of less control rather than more, since appropriate
variation in the lower-scale variables is often the means by which the
higher-scale relation is kept in place.

The second is that not every intervention is equally probative. The intervention
has to test the relation through the relevant pathway rather than destroy or
bypass it. Perturbations to workload or demand can reveal the controller's
response, while clamping the regulated variable itself can sever the very
feedback that constitutes the control. Stillness under a clamp doesn't show that
nothing was being controlled.

For the purposes of this paper, corrective control has four evidential marks. The
evidence should include a perturbation that threatens the projectible profile, a
variable or relation to which the response pathway is dynamically coupled, a
response that changes lower-scale conditions, and a result in which that response
preserves a higher-scale relation supporting projection. Feedback on its own
doesn't satisfy these conditions, because the evidence has to show that the
feedback explains why the profile remains projectible under perturbation.

That evidence is a good deal more specific than the claim that a property cluster
is stable, or that its properties stand in some causal order. It marks the most
demanding diagnostic question: not just whether the profile persists, and not just
whether it is ordered, but whether a specified corrective pathway preserves the
relation that makes the profile projectible.

\section{The diagnostic grid}

The diagnostic grid is a set of evidential questions rather than an exhaustive
taxonomy or a guaranteed sequence. Control-secured cases will usually display
stability, network order, and maintenance as well, but the inference doesn't run
in reverse. A stable cluster needn't have an internally ordered profile, a
network-ordered kind needn't have a specified maintainer, and a maintained kind
needn't be maintained by corrective control.

The grid also has off-diagonal cases. A bureaucratic status label may be
maintained by forms and offices while its associated properties remain weakly
ordered; a transient machine-failure profile may have ordered causal dependencies
without any stabilizer that keeps it in place.

The grid isolates five questions that are recurrently compressed in HPC
discussions: what gets projected, what profile supports the projection, what order
makes the profile non-accidental, what maintains the profile, and whether that
maintenance is corrective control. Selection, drift, institutional scaffolding,
exemplar effects, and path dependence may each stabilize projectible profiles in
ways that deserve separate treatment. These cases matter because they may secure
projectibility without fitting neatly into the strict control role; the grid is
meant to prevent conflation, not to classify every stabilizing process.

This is the constructive response to Onishi and Serpico's dilemma. Slater-style
cliquish stability, Khalidi-style causal-network order, Boyd-style maintenance,
and control-theoretic homeostasis are separated because no one of them should be
made to carry all the weight of natural-kind realism.

\begin{table}[H]
\centering
\small
\begin{tabular}{>{\raggedright\arraybackslash}p{0.21\linewidth}
                >{\raggedright\arraybackslash}p{0.35\linewidth}
                >{\raggedright\arraybackslash}p{0.30\linewidth}}
\toprule
Diagnostic demand & Evidence needed & Failure mode \\
\midrule
Projection target & A specified field-relative inference, prediction,
explanation, or intervention the category is supposed to support & Kind talk
without saying what the category lets us project \\
Stable profile & A Slater-style projection from observed subcluster to
unobserved cluster features or behavior across relevant contexts, plus invariance
evidence for explanatory or interventional payoffs & Mere recurrence without
projectible payoff \\
Network order & An ordering relation that explains which projections are
licensed, such as core-to-derivative, conditioning, licensing, or
causal-normative dependence & Causal or ordering vocabulary used for an unordered
list or mere cue \\
Maintenance & A stabilizer whose removal would make the profile fragment, drift,
or dissolve & Treating persistence as if a maintainer had been specified \\
Corrective control & Perturbation-sensitive correction that preserves a
higher-scale relation through lower-scale variation & Treating stability,
maintenance, or feedback as control \\
\bottomrule
\end{tabular}
\caption{A diagnostic grid for projectible profiles and homeostatic control.}
\label{tab:diagnostics}
\end{table}

The grid is meant to turn the question around. It asks the analyst to name the
intended projection first and then justify each stronger securing claim, ending
only where a maintainer corrects perturbations. The aim is to make the evidential
demand behind \term{homeostatic} explicit.

Nothing in the argument requires every historical use of \term{homeostatic} in
HPC theory to have meant feedback control. The next step is to test separability
in cases: stable profile, network order, maintenance, and corrective control
should answer different evidential questions and come apart in practice, not just
in the table.

\section{The cases}

The cases are arranged by evidential role. Their taxonomic standing differs: the
Watt governor calibrates the control architecture; glycemic regulation supplies a
biological profile plausibly secured by such architecture; interjection tokens show
projectibility and maintenance without a specified controller; and monitored
agreement production shows what a linguistic dependency would have to add to count
as control-secured in an online repair loop. Together, the cases test whether a profile
supports projection, what order explains that projection, what maintains the
profile, and whether the maintainer is corrective control.

The spread across machine, biological, and linguistic patterns is part of the
test. The diagnostic concerns how projectibility is secured, a question not
internal to any one science. If the same grid discriminates among the governor,
glycemic regulation, interjection tokens, and monitored agreement production, the
distinction belongs to the theory of kinds rather than the folklore of a
particular field.

\subsection{Calibration case: Watt governor}

The Watt governor is a useful calibration case because its control structure is
explicit. The governor-engine assembly is the system; flyballs, linkage, and valve
relations are controller components; shaft speed is the lower-scale feedback
variable; and the higher-scale controlled relation matches steam admission to
workload. A change in engine speed moves the governor's flyballs; their position
changes the valve aperture; the aperture changes the flow of steam; and the steam
flow changes the speed of the shaft. A clamp placed directly on shaft speed would
sever the incoming causal relations and destroy the control relation altogether
\citep{weinberger2026HomeostasisCausalControl}.

Direct clamping still leaves ordinary perturbation tests available. A change in
workload or demand is the right kind of perturbation, because it lets the feedback
relation operate and reveals whether component-level adjustment preserves the
higher-scale matching relation.

The governor shows why \term{homeostatic} shouldn't mean constant. Its
component-level variables are allowed to vary precisely so that steam admission
matches workload across changes in speed. That pattern calibrates the strict
control demand: a perturbation is met by responsive adjustment, and a
higher-scale relation is preserved through that adjustment.

The governor's role is calibrational. As an artifact with a target fixed by its
designer, it displays control-system predictability for comparison with Boyd-style
kind projectibility. Candidate kind claims can then be tested by asking whether an
analogous perturbation-sensitive architecture secures their projectible profiles.
The governor doesn't calibrate everything that matters, since its target is fixed
externally. It calibrates lower-scale feedback producing higher-scale matching;
the next case supplies the harder lesson about how a range can be constituted
from within the system.

\subsection{Biological extension: glycemic regulation}

Glycemic regulation is included because it shows why the control diagnostic should
be relation-preserving rather than set-point-preserving. The relevant candidate is
the normoglycemic regulatory profile of an organism, not glucose as a substance or
a single blood value. The regulated variable is blood glucose concentration; the
maintaining architecture is the organism's endocrine and metabolic organization;
and the projectible profile consists in coordinated responses across feeding,
fasting, activity, and other perturbations.

The standard feedback representation already gives the case a control-like
architecture. It includes a device that sets a normal range, a sensor that
measures the regulated variable, a detector that compares the measured value
against the range and emits an error signal, a controller, and an effector that
moves the variable back toward the range \citep{bichMossioSoto2020}. Taken at
face value, this is why the case looks control-secured.

The projectible payoff is the organism's metabolic response across changing
conditions. Given the normoglycemic regulatory profile, one expects coordinated
changes in insulin, glucagon, hepatic output, and glucose uptake across feeding,
fasting, and activity, rather than an isolated reading that happens to fall within
a range.

The standard representation already reserves the word \term{controller} for the
component that reads the error signal and drives the effector. The vocabulary of
the calibration case and of the biological case coincides, and not by accident,
since both are feedback architectures in which a sensed departure drives a
correcting response. The diagram doesn't vindicate the word; the architecture
has to earn it.

Two complications make that question sharper. The first concerns the set point,
since population correlations for glucose and calcium may fit an equilibrium model
better than they fit a defended set point \citep{fitzgeraldBean2018}.

The second is that \textcite{bichMossioSoto2020} argue that the feedback-loop
diagram is explanatorily thin: it localizes functions too neatly, represents a
hierarchical organism as a flat component chain, narrows the explanation to one
regulated variable, and assumes a normal range without explaining what
establishes it.

They move instead to organizational closure, on which glucose homeostasis is a
manifestation of organismic self-maintenance and energy-resource management, not
just the tracking of an externally given number. Read this way, the corrective
structure is real, but the set point is the element the cybernetic analogy gets
wrong.

That move marks the real contrast with the governor. The governor's target is set
from outside by a designer, whereas the organism's range has to be explained from
within the organized system. The higher-scale relation that glycemic control
preserves is a relation between glucose availability and the organism's continued
integrated functioning, held in place through wide variation in component-level
variables across feeding, fasting, and activity. A fixed reading would misdescribe
that relation. A biological case counts as control-secured only when the evidence
displays this organized correction; mere range maintenance is insufficient.

Glycemic regulation is the biological stress test for the strict control
diagnostic: the stable profile is maintained by organized responses to
perturbation, and what's preserved is best understood as the relation rather than
the reading. It also shows where the governor case stops: lower-scale feedback
matters, but the biological target must be explained as part of organismic
self-maintenance.

\subsection{Contrast case: interjection tokens}

English interjection tokens are the contrast linguistic case. The unit is
token-level: a form counts here only when it occurs with the interjection profile
of prosodic detachability, weak syntactic integration, interactional force, and
uptake-relevance.\footnote{For the broader claim that word-kinds can be understood
as homeostatic property clusters, see \textcite{miller2021}. This case is
narrower: it asks which evidential demands interjection tokens and form-sequence
constructions satisfy.} A detached response cry \mention{shit} is an interjection
token, while \mention{shit} in an ordinary nominal environment is a noun token.

The case has two levels of projection. At the broad category level, an
interjection token licenses weak expectations about prosodic detachability, low
syntactic integration, non-inflection in the relevant occurrence, interactional
force, and uptake-relevance \citep{ameka1992,norrick2009}. Observing some of
these properties licenses expectations about the others, which makes the profile
projectible rather than a mere list.

Stronger projections arise at the lower constructional level. Forms such as
\mention{huh?}, \mention{mmhm}, and \mention{oh} support expectations about
sequential position and interactional uptake, including repair initiation,
continuation, and a change in epistemic state \citep{dingemanse2024}, while forms
such as \mention{wow}, \mention{oops}, \mention{ouch}, \mention{yuck},
\mention{shit}, \mention{hey}, \mention{well}, and \mention{uh} invite
expectations about stance, event management, attention, transition, or delay.
\textcite{norrick2009} is helpful here because he treats interjections as an
open-ended category with recurrent pragmatic-marker functions. The two-level
structure matters because the diagnostic should not force projectibility to be all
or nothing.

The contrast between \mention{oh} and \mention{uh} shows the kind of projection at
issue. In an appropriate sequential position, \mention{oh} leads us to expect a
change-of-state uptake: the speaker has just registered, noticed, or reclassified
something. \mention{uh}, by contrast, leads us to expect delay or floor-holding
rather than a completed uptake. These are genuine projections from form and
placement to interactional role, but they don't yet show a controller preserving
the use profile.

The list is heterogeneous, since several of these forms are analysed as discourse
markers, fillers, summonses, or response tokens rather than as central
interjections. This heterogeneity is part of the test, because the contrast case
is a projectible profile of interjection tokens with kindhood partly at issue,
rather than a classical kind with settled lexical boundaries.

The likely stabilizers include conventionalization, turn-taking routines, prosodic
recognizability, acquisition, register, recurrent response patterns, and
metapragmatic learning, and together these make the profile stable and useful. They
also give parts of the profile some network order. The relevant order has to be
asymmetric rather than mere non-independence: interactional autonomy conditions
prosodic detachability and syntactic non-integration, sequential position
conditions available uptake, and prosodic contour helps disambiguate the uptake at
stake. That gives interjection tokens projectibility, some network order, and
plausible stabilizers. It doesn't yet give them corrective control.

In this case, maintenance is best sought in recurrent interactional practices that
keep form, prosody, sequential position, and expected uptake aligned across
speakers. These practices operate at several levels. Conventionalization and
community transmission entrench form--function pairings over time; acquisition,
modelling, and metapragmatic teaching stabilize use across speakers; local repair
handles occasional trouble in a conversation. Removing the recurrent practices
wouldn't produce a single momentary error so much as drift in the projectible
profile itself.

The homeostasis question is more demanding. For interjection tokens, the
higher-scale relation would be a use-constructional relation linking turn
position, prosodic packaging, syntactic detachment, and expected uptake. The
lower-scale variables would include token form, intonation, duration, sequential
placement, phonological reduction, and degree of syntactic detachment.

Given the token-level unit, relevant perturbations have to stay inside an
otherwise interjectional occurrence. Ordinary category shifts, such as
\mention{wow} becoming an attributive modifier or \mention{shit} appearing as an
ordinary noun, move to a different construction. Better candidates are anomalous
prosody, misplaced sequential deployment, over-integration in an otherwise
interjection token, or an ambiguous token that receives repairable uptake. Failure
of uptake is an outcome or a trigger for repair, rather than the perturbation
itself.

A control claim would have to identify one loop rather than move among several
possibilities. In an online interactional loop, failed or ambiguous uptake could
lead to clarification or repair that restores sequential intelligibility. In a
developmental loop, learner misuse could lead to correction or modelling and
convergence on adult use. In a community-transmission loop, repeated uptake
patterns could entrench a form--function pairing. These are neighbouring
stabilizers, not yet one control architecture.

The contrast with monitored agreement production turns on evidential
tractability, not on a special standard for interjections. The online interjection
loop may use the same broad parsing-and-monitoring capacities used in speech
repair. Agreement supplies a sharply specified relation, controller features to
target forms, and an independently characterized perturbation. Interjection uptake
offers a more diffuse relation and perturbation classes, such as anomalous
prosody, misplaced sequencing, or over-integration, without an equivalent
paradigm. Until the relevant loop is specified, interjection tokens remain stable,
partly network ordered, or maintained without yet being control-secured.

\subsection{Boundary case: monitored agreement production}

Monitored agreement production is the boundary case where the homeostatic label
comes closest to being earned. The candidate is the monitored production process
preserving an utterance-level agreement relation, not the type-level grammatical
regularity by itself.

In agreement systems, agreement is systematic covariance between controller and
target: one expression fixes features that other expressions must express.
\olang{I go} is the canonical English pairing, while \ungram{\olang{I goes}}
mismatches the first-person controller with a third-person verb form. In Corbett's
terms, the agreeing expressions are targets, and canonical agreement is the case
in which the features take matching rather than mismatching values
\citep{corbett2006}. The features may be number, person, gender, class
membership, or case. The case concerns monitoring across agreement features, with
gender as one possible feature.

That terminology is useful but not decisive. Corbett's controller--target relation
identifies a dependency structure in agreement. By itself, that relation doesn't
amount to a feedback-control architecture. The controller doesn't sense a
mismatch, compare target forms with a standard, or issue a corrective signal.
Agreement morphology establishes the relation whose preservation matters;
production monitoring and self-repair are where a control-theoretic claim has to
look for feedback.

Agreement has a straightforward projectible payoff. Once the controller's features
and the agreement domain are known, target forms can be projected across the
domain. Number agreement lets us project verb form from subject number. Noun-class
agreement lets us project class-marking across verbs, adjectives, and demonstratives
\citep{ashton1947,corbett2006}. The dependency is also ordered rather than merely
stable, because the controller--target relation is asymmetric: the controller
supplies the feature values that agreement targets must match, even when the
controller also bears or expresses those values.

For the grammatical dependency, the first evidential demands are straightforward:
agreement has a projection target, a stable profile, and an ordered dependency
relation. The harder question is whether any maintenance process for that relation
is also corrective control.

This is where scale matters. Agreement as a type-level grammatical regularity may
be stable and ordered, but online self-monitoring operates on particular
utterances. Repairing \olang{The key to the cabinets are} to \olang{The key to
the cabinets is} preserves this utterance's agreement relation. The case is
therefore a boundary case: token-level control evidence is possible without
settling control of the broader grammatical kind.

The control question arises only when that dependency is embedded in a corrective
process. Agreement attraction helps isolate one perturbation: plural local nouns
reliably increase agreement errors, while animacy and length do not
\citep{bockMiller1991}. The plural attractor perturbs the implementation of
subject--verb number agreement; the agreement error is the failure state.

Levelt's repair model supports that loop only in a limited way. Speakers monitor
inner or overt speech, interrupt, and repair \citep{levelt1983}; this is
parsing-based monitoring, not direct access to production machinery. The monitor
is not agreement-specific and may also matter for interjection uptake. Agreement
is cleaner because the monitored relation and perturbation can be specified
independently: utterance-level agreement, target forms, and repair or recasting.

The agreement-specific character of the repair comes from the projected dependency
itself. Once the controller's features and the domain are fixed, target forms are
expected to express those values, so repair restores the higher-scale relation that
made the profile projectible rather than only smoothing local fluency. The repair
counts as agreement-specific only when the corrected form is selected because it
restores the controller--target relation.

A simple attraction case makes the loop concrete:

\ea\label{ex:agreement-repair}
\olang{The key to the cabinets are~-- sorry, is~-- on the table.}
\z

The perturbing cue is the plural attractor \mention{cabinets}; the threatened
relation is agreement with \mention{key}; the production-level variable is the verb
form; and the response pathway is repair from \mention{are} to \mention{is}. The
repair restores the dependency relation that licensed the projection from
controller to target in the first place.

This keeps constraint and correction apart. A grammatical description may mark a
mismatch as ill-formed without thereby supplying a corrective pathway. Corrective
control requires the additional process by which a departure is detected and
answered. The four marks may be met in the online repair loop: the mismatch
threatens the projectible profile, agreement is the tracked relation, repair
changes the target forms, and the higher-scale relation is preserved through
that change.

Acquisition and normative correction are different candidate loops; coordination
would be needed to make them parts of a single system. In acquisition, learners
converge on agreement patterns over developmental time; in normative contexts,
teachers, editors, or interlocutors may reject or correct mismatches. These
processes help stabilize agreement and may provide further control-like evidence,
but they operate at different scales from online monitoring. The honest conclusion
is conditional: monitored agreement production supplies a specified linguistic
loop, which avoids a loose analogy from grammatical dependency to feedback control,
and the homeostatic label depends on that loop doing the relevant corrective work.

Grammaticality at large is a harder and more contested candidate.\footnote{For the
broader grammaticality-as-kind project, see
\textcite{reynoldsmiller2026grammaticality}. This paper only uses monitored
agreement production to show what a specified linguistic control loop would have to
look like.} If correction, repair, acquisition, transmission, and uptake together
formed an architecture that returned usage to a community-sensitive profile,
grammaticality could turn out to be control-secured, but whether grammaticality is
a single kind, and which loops would be doing the controlling, remains in dispute.

The diagnostic leaves the possibility open without granting it, and monitored
agreement production shows how a control-secured linguistic case would have to
specify its loop.

\section{What the diagnostic changes for HPC theory}

The grid matters for HPC-style applications in linguistics and social ontology.
Linguistic categories can have projection targets, stable profiles, network order,
and maintainers without all of them being homeostatic in the strict
control-theoretic sense, and social practices can be stabilized by records, offices,
sanctions, training, and uptake without thereby constituting single homeostatic
control systems.

Social-status assignments are a clear case in point: they may support projection
through causal-normative networks of uptake, recognition, transition, and
constitutive ties without thereby carrying moral warrant
\citep{reynolds2026effective}. The diagnostic asks each case which evidential
demands it satisfies before control is inferred from the mere fact of persistence.

A natural objection is that social correction and metalinguistic policing are
themselves forms of feedback, and sometimes they do function that way. The
concession strengthens the diagnostic because it means that some linguistic and
social categories may indeed be control-secured, but only where the loop is
specified at the right scale.

The monitored agreement case keeps that condition visible: its candidate
controller is the monitored production loop preserving an utterance-level
agreement relation. Broader categories would need their own coordinated
architecture of correction, repair, acquisition, transmission, and uptake. The
evidence has to earn the upgrade case by case; the label can't be granted on the
strength of stability alone.

SPACER gives a useful proof of concept for this discipline. The dataset pairs
naturalistic single-word substitution errors from Switchboard with speaker
self-repairs and with comprehenders' offline text-editing corrections
\citep{upadhyeEtAl2025spacer}. The perturbation is a lexical substitution; the
candidate tracked relations include lexical fit, semantic distance, phonemic
distance, and contextual predictability; and the response pathways are speaker
repair and comprehender replacement. That structure lets the diagnostic become an
empirical comparison without assuming that all correction belongs to one loop.

Here SPACER matters methodologically. A dataset of this structure lets one ask
whether the same perturbation measures predict speaker-side and comprehender-side
correction, or whether apparent recovery of intelligibility is maintained by
neighbouring stabilizers. Until the same tracked relation and response pathway are
specified, speaker monitoring, comprehender editing, and intelligibility recovery
shouldn't be folded into one homeostatic architecture.

This is also why target-declaration can't be gamed by relabelling. If speaker
repair and comprehender editing show different relations to the perturbation
measures, redescribing both as recovery of intelligibility doesn't make them one
control loop. The measured contrast disciplines the target claim.

This constraint governs the upgrade from a declared projection target to a control
claim. It doesn't choose the initial target from all possible descriptions of the
profile; that choice still has to be justified by the projection, explanation, or
intervention the category is meant to support.

The same distinction bears on the more formal side of HPC work. One temptation is
to operationalize homeostasis as a return-to-basin claim, asking whether a
variable comes back to a set point after perturbation. The governor and glycemic
regulation both show why that test is too narrow, since what each system preserves is
a matching relation: workload to steam flow in the one case, and glucose
availability to self-maintenance in the other. In both, the relation is held in
place through variation in the lower-scale variables.

A test that measures only the return of a variable to a fixed value can miss
control that works precisely by allowing the variable to move. The basin is better
defined over the relation the system preserves than over any set of fixed values.

For HPC-style work in linguistics this implies a division of labour. The
home-field result is sober: negative or conditional, not a failed positive case.
Interjection tokens show how far stable pragmatic projectibility can go without a
specified controller, and monitored agreement production shows what a local
control-secured loop can look like. Many categories that an HPC analysis might be
tempted to call homeostatic are, on inspection, stable projectible profiles,
maintained profiles, or network-ordered kinds, which are real and worth a
realist's defence without being control-secured.

What philosophers of science gain is a domain-neutral test for those stronger
claims. The diagnostic should also obey its own discipline: given a category
already offered as a Boyd-style projectible kind, it specifies what additional
evidence would be needed before each upgrade is licensed.

It's narrower than a general test for natural kindness, and it doesn't claim that
all projectible categories share one stabilizing structure. This reflexive check
is Boydian in spirit: Boyd treats the category \term{natural kind} itself as a
natural kind in metaphysics and epistemology \citep[168]{boyd1999}.

Its usefulness depends on whether it blocks overextension in cases like
interjections while still
recognizing control-secured cases, such as glycemic regulation or the monitored
production loop that preserves an utterance-level agreement relation, when the
relevant loop is specified.

\section{Conclusion}

Projectibility is the point throughout. The finding is separability: stable
profile, network order, maintenance, and corrective control are distinct
evidential achievements that can come apart. A stable cluster can
underwrite projection without being control-secured, because the reliability of
the profile for a specified inferential purpose is not yet evidence of corrective
organization under perturbation. Reserving \term{homeostatic} for corrective
control is the recommendation that follows from that result, not the premise that
produces it.

\section*{Data and code availability}

The \href{https://brettrey.github.io/not-every-stable-cluster-is-homeostatic/}{companion
SPACER analysis} is available online, and the source code is available in the
\href{https://github.com/BrettRey/not-every-stable-cluster-is-homeostatic/tree/main/analysis/spacer-diagnostic}{project
repository}. The workflow downloads the public SPACER data from the project
repository associated with \textcite{upadhyeEtAl2025spacer}; the raw SPACER data
are not redistributed here.

\section*{Acknowledgements}

Drafting and revision benefited from OpenAI Codex (GPT-5), GPT-5.5 Pro, and Claude
Opus 4.8. All content has been reviewed and revised by the author, who takes full
responsibility for the final text.

\clearpage

\printbibliography

\end{document}
