---
title: "Fake Data Recovery"
---

```{r}
source("R/setup.R")
shared <- simulate_correction_profiles(shared_profile = TRUE)
distinct <- simulate_correction_profiles(shared_profile = FALSE)
fit_shared <- fit_fake_recovery(shared)
fit_distinct <- fit_fake_recovery(distinct)
```

Fake-data simulation checks whether the proposed model can recover the distinction that matters before it is pointed at SPACER.

## Shared Profile Scenario

In the shared-profile simulation, the process-by-predictor interactions are set to zero.

```{r}
shared_terms <- summary(fit_shared)$coefficients |>
  as.data.frame() |>
  tibble::rownames_to_column("term") |>
  dplyr::filter(stringr::str_detect(term, "processspeaker:"))
knitr::kable(shared_terms, digits = 3)
```

## Distinct Stabilizer Scenario

In the distinct-profile simulation, speaker and comprehender correction differ in their phonemic-distance and surprisal slopes.

```{r}
distinct_terms <- summary(fit_distinct)$coefficients |>
  as.data.frame() |>
  tibble::rownames_to_column("term") |>
  dplyr::filter(stringr::str_detect(term, "processspeaker:"))
knitr::kable(distinct_terms, digits = 3)
```

## Check

The fake-data exercise is not evidence about SPACER. Its role is procedural: it checks that the model target is capable of distinguishing "same correction profile" from "different correction profile" under known conditions.
