---
title: "Fit the Joint Model"
params:
  refresh_model: false
  fit_bayes: false
---

```{r}
source("R/setup.R")
metrics <- load_spacer_metrics()
dat <- make_process_data(metrics)
predictors <- make_predictor_set()
model <- fit_or_load_joint_glmer(dat, predictors = predictors, refresh = params$refresh_model)
```

The default fit is a joint logistic mixed model:

```{r}
formula(model)
```

The process-by-predictor interactions are the main quantities. They test whether a predictor has a different association with speaker repair than with comprehender correction.

## Coefficients

```{r}
terms <- tidy_glmer(model)
knitr::kable(terms, digits = 3)
```

## Process Contrasts

```{r}
interaction_terms <- terms |>
  dplyr::filter(stringr::str_detect(term, "processspeaker:")) |>
  dplyr::arrange(dplyr::desc(abs(estimate)))
knitr::kable(interaction_terms, digits = 3)
```

## Prediction Curves

These are population-level predictions with random effects held out.

```{r}
phonemic_grid <- make_prediction_grid(model, "phonemic_z", predictors)
plot_prediction_grid(phonemic_grid, "Phonemic distance (z)")
```

```{r}
surprisal_grid <- make_prediction_grid(model, "surprisal_z", predictors)
plot_prediction_grid(surprisal_grid, "Target minus error surprisal (z)")
```

## Bayesian Fit

The Bayesian model is the preferred final analysis if the workflow becomes a paper appendix or companion repository. It is off by default because it requires a working Stan backend and takes longer to render.

```{r}
#| eval: !expr params$fit_bayes
bayes_model <- fit_joint_brms(
  dat,
  predictors = predictors,
  chains = 4,
  iter = 2000,
  cores = 4,
  backend = "cmdstanr"
)
bayes_model
```

The priors are weakly informative on the log-odds scale. Before using the Bayesian fit for publication, add prior predictive checks and compare posterior predictive summaries against the observed correction rates.
