---
title: "Model Checks"
params:
  refresh_model: false
---

```{r}
source("R/setup.R")
metrics <- load_spacer_metrics()
dat <- make_process_data(metrics)
predictors <- make_predictor_set()
model <- fit_or_load_joint_glmer(dat, predictors = predictors, refresh = params$refresh_model)
```

The model should reproduce basic features of the data before its process contrasts are interpreted.

## Observed and Fitted Rates

```{r}
checks <- observed_vs_predicted(model, dat)
knitr::kable(checks, digits = 3)
```

```{r}
plot_observed_predicted(checks)
```

## Simulated Correction Rates

The following check simulates replicated outcomes from the fitted mixed model and compares process-level rates with the observed rates.

```{r}
sim_rates <- simulate_process_rates(model, nsim = 100)
plot_simulated_rates(sim_rates, dat)
```

## What This Does Not Check

These checks are intentionally modest. They do not validate the lexical metrics, the offline comprehender task, or a causal interpretation of the slopes. They only ask whether the fitted model can reproduce broad correction rates before the interactions are used in the ladder report.
