---
title: "Multiverse"
---

```{r}
source("R/setup.R")
metrics <- load_spacer_metrics()
```

A small multiverse exposes reasonable analytic choices. The default version uses fast logistic models without random effects as a screening layer. Variants that change the philosophical conclusion should be refit with the joint mixed model.

```{r}
variants <- tidyr::crossing(
  phonemic_measure = c("ipa", "edit"),
  include_semantic = c(TRUE, FALSE),
  include_frequency = c(TRUE, FALSE),
  include_surprisal = c(TRUE, FALSE)
) |>
  dplyr::filter(include_semantic | include_frequency | include_surprisal)

fit_variant <- function(phonemic_measure, include_semantic, include_frequency, include_surprisal) {
  dat <- make_process_data(metrics, phonemic_measure = phonemic_measure)
  predictors <- make_predictor_set(
    include_semantic = include_semantic,
    include_phonemic = TRUE,
    include_frequency = include_frequency,
    include_surprisal = include_surprisal,
    include_length = FALSE
  )
  f <- stats::as.formula(
    paste("corrected ~ process * (", paste(predictors, collapse = " + "), ")")
  )
  model <- stats::glm(f, data = dat, family = stats::binomial())
  summary(model)$coefficients |>
    as.data.frame() |>
    tibble::rownames_to_column("term") |>
    dplyr::filter(stringr::str_detect(term, "processspeaker:")) |>
    dplyr::transmute(term, estimate = Estimate, std_error = `Std. Error`)
}

multiverse <- variants |>
  purrr::pmap_dfr(function(...) {
    args <- list(...)
    fit_variant(...) |>
      dplyr::mutate(
        phonemic_measure = args$phonemic_measure,
        include_semantic = args$include_semantic,
        include_frequency = args$include_frequency,
        include_surprisal = args$include_surprisal
      )
  })
```

## Interaction Estimates Across Variants

```{r}
knitr::kable(multiverse, digits = 3)
```

```{r}
ggplot2::ggplot(multiverse, ggplot2::aes(x = estimate, y = term, color = phonemic_measure)) +
  ggplot2::geom_vline(xintercept = 0, linetype = "dashed") +
  ggplot2::geom_point(alpha = 0.75) +
  ggplot2::labs(x = "Process interaction estimate", y = NULL, color = "Phonemic measure")
```

## Interpretation Rule

The multiverse is not a pile of robustness badges. Its job is to identify whether the ladder conclusion depends on fragile coding choices. If the process-by-phonemic-distance contrast persists across reasonable variants, it is a strong candidate for the report. If a contrast appears only under one coding path, it should be treated as exploratory.
