---
title: "Ladder Report"
params:
  refresh_model: false
---

```{r}
source("R/setup.R")
metrics <- load_spacer_metrics()
dat <- make_process_data(metrics)
predictors <- make_predictor_set()
model <- fit_or_load_joint_glmer(dat, predictors = predictors, refresh = params$refresh_model)
terms <- tidy_glmer(model)
```

## Result Template

```{r}
knitr::kable(ladder_result_template())
```

## Current Model Signal

```{r}
signal <- terms |>
  dplyr::filter(stringr::str_detect(term, "processspeaker:")) |>
  dplyr::mutate(
    direction = dplyr::case_when(
      estimate > 0 ~ "larger speaker slope",
      estimate < 0 ~ "smaller speaker slope",
      TRUE ~ "no contrast"
    )
  ) |>
  dplyr::arrange(dplyr::desc(abs(estimate)))
knitr::kable(signal, digits = 3)
```

## Provisional Warrant

On the default SPACER metric surface, the workflow should be read as a test of profile alignment. A large process-by-predictor contrast counts against the simple story that speaker self-repair and offline comprehender correction instantiate one shared correction profile.

That result does not show that the speaker-side process is homeostatic. Nor does it show that the comprehender-side task is a controller. The stronger philosophical use is negative and diagnostic: SPACER makes it possible to distinguish recovery of intelligibility from evidence for a single corrective-control architecture.

## Ladder Placement

```{r}
placement <- tibble::tribble(
  ~rung, ~status, ~reason,
  "Projection target", "Met", "The workflow declares a profile-comparison target before modeling.",
  "Stable profile", "Testable", "Repeated items and responses allow correction profiles to be estimated.",
  "Network order", "Partial", "Predictor ordering is plausible but not established as causal order.",
  "Maintenance", "Partial", "Speaker repair and comprehender correction are stabilizing pathways, but they are not one pathway.",
  "Corrective control", "Not warranted from SPACER alone", "The dataset supports profile comparison, not direct evidence for one online controller."
)
knitr::kable(placement)
```

## Paper Use

The cleanest use in the paper is as an empirical extension or appendix:

> SPACER shows how the ladder changes the empirical question. The issue is not whether correction occurs, but whether the same perturbation measures predict speaker self-repair and comprehender correction in the same way. Divergence supports the diagnostic caution that intelligible recovery can be maintained by neighboring stabilizers rather than by one specified corrective-control system.
